Distribution-oblivious Programming with Dish

### 2019-11-20 pldi/draft.tex
Update overview

@@

Distributability classes, based on how

>>>>

As outlined earlier, the key insight behind \sys is that the shell already describes parallel computations over streams.
Intuitively,
 (i) the parallel-composition operator exposes task parallelism,
 (ii) commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
 (iii) the pipe operator enables both task and data parallelism, and
 (iv)  other composition operators can be viewed as barrier candidates.
To leverage these insights and correctly distribute a pipeline such as $p_1$, \sys must solve several challenges.

Understanding the shell's primitives is not enough;
  \sys has to understand the distribution characteristics of individual commands used to compose larger programs.
This challenge can be broken down into two parts.
First, there is a need to understand the commands already in the shell---that is, built-ins that come with any shell and support its core functions.
One example is the set of commands defined by the \textsc{POSIX} standard~\cite{}.
\sys addresses this part by a careful study of a group of commands, identifying a small but clear set of well-understood distributability classes.
This set is already useful enough to allow the composition of practical pipelines~\sx{eval}.

The second part is expressing such characteristics extensions that fall outside the built-in set, allowing new commands to leverage \sys's power.
Commands are arbitrary programs, but still conform to the shell's pipeline interface.
To solve this part, \sys leverages the prior study to define a small but expressive domain-specific language (DSL) for describing a command's key properties with respect to its distributability.
Using this DSL, developers of commands (not users of the shell) can quickly and easily capture the class of the command.
A command developer can only annotate the default behavior;
  if a new flag is added a few years later, and if that flag alters the command's class, only then a new annotation is needed.

Given a set of commands composed together by the shell primitives, \sys's next challenge is to analyze it and identify candidate subexpressions for distribution.
\sys addresses this challenge with a rewriting \todo{pass} that performs dataflow analysis identifying maximal distributability sub-expressions.
It converts the AST to a dataflow graph, and runs...
This analysis is not trivial, as many of the shell's dynamic features have to be taken into consideration---examples include shell expansion, environment variables, file input sizes, \etc

Finally, given a distributed pipeline with runtime \todo{annotations}, \sys needs to schedule and execute it.
This amounts to solving a few challenges, achieved by \sys's runtime component, two of which stand out.
First, \sys needs to wrap single-node commands with corrective runtime interposition based on their classes;
  this is achieved by thin wrappers for all the commands involved in the program, generated just-in-time using the distributability specification outlined earlier.
The reasoning behind wrappers being thin is to not interfere with legacy functionality---commands have many flags accumulated over long periods of time.

Second, \sys needs to fill-in planing details left blank by the rewriting pass.
This is achieved with \emph{planner activators}, commands that yield (and pass their arguments) to the scheduler.
From the point of view of the shell, activators are higher-order commands \'{a} la \ttt{xargs} and \ttt{time} that take other commands as arguments---a technique that achieves the latest possible binding, when critical runtime information is available.

The next few sections discuss the details.
They also outline other, less obvious challenges---such as the synchronization of environment variables, distributed file operation, \etc.

\section{Distributability Classes}

@@

\begin{table}[h]
\center
\footnotesize
\setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Distributability Classes}.
    Broadly, \unix commands can be broken down into to five classses.
  }
}
\begin{tabular}{l @{\extracolsep{\fill}} lll}
\toprule
Class                           & Example Commands        & POSIX            & Core-utils\\
\midrule
Stateless                       &                         &                  &           \\
Pure                            &                         &                  &           \\
(Distr.) File-System            &                         &                  &           \\
Side-effectful                  &                         &                  &           \\
Irreversibly Side-effectful     &                         &                  &           \\
% \midrule
% Shell PL Constructs             &                         &                  \\
% \etc                            &                         &                  \\
\bottomrule
\end{tabular}
\label{tab:classes}
\end{table}

Define distributability 
Explain distributability classes, based on how
